---
title: "User-similarity-based recommendation"
---

&nbsp;

A basic strategy for recommendation is, find users similar to the new user, and
recommend the new user with what similar users like. Based on this strategy, a
technique, collaborative filtering, was invented to making predictions (filtering) 
about the interests of a user by collecting preferences or taste information from 
many similar users (collaborating).

&nbsp;

Here we implemented a user similarity based collaborating filtering algorithm to 
predict ratings, and this algorithm was further applied to our 
[shiny app](https://hzhong.shinyapps.io/movierecommender/).

&nbsp;

### Dependencies

```{r, message=FALSE, class.source = "fold-show"}
library(tidyverse)
```

&nbsp;

# Data preparation

```{r, warning=FALSE, message=FALSE}
path = "./data/small/ratings.csv"

ratings_tidy = read_csv(path) %>% 
  janitor::clean_names() %>% 
  select(-timestamp)

ratings = ratings_tidy %>% 
  nest(movie_id, rating)
```

In the data preparation process, we read the raw rating file and nested the 
two column `movie_id` and `rating` for following analyses' convenience.

&nbsp;

# Algorithm

The user similarity based recommendation algorithm predict ratings with the following
procedures:

- Find $X$ most similar users to the given user (Here we set $X$ to 10)
- Find all movies that these similar users have rated and the given user has not watched
- Predict the given user's ratings on these movies based on the similarity measure and ratings of similar user's

&nbsp;

### Measure of similarity

We built two functions, `calc_pc_sim`, `calc_cos_sim`, to calculate the similarity
between users. In subsequent analysis, we will compare the performance of these
two measure and choose the better one.

&nbsp;

##### Measure by Pearson correlation

Let's denote $w_{uv}$ as the similarity between two users($u$, $v$), $r$ as movie
ratings, $m$ as movie, then $w_{uv}$ can be quantitatively defined by Pearson 
correlation:

$$
w_{u v}=\frac{\displaystyle \sum_{m \in M}\left(r_{u m}-\bar{r}_{u}\right) \cdot\left(r_{v m}-\bar{r}_{v}\right)}{\displaystyle \sqrt{\sum_{m \in M}\left(r_{u m}-\bar{r}_{u}\right)^{2} \sum_{m \in M}\left(r_{v m}-\bar{r}_{v}\right)^{2}}}
$$

```{r}
calc_pc_sim = function(u_data, v_data){
  
  calc_df = u_data %>% 
    inner_join(v_data, by = "movie_id", suffix = c("_u", "_v"))
  
  # two users should at least watched 3 same movie
  if (nrow(calc_df) < 3) {
    return(-1)
  }
  
  r_mean_u = u_data %>% pull(rating) %>% mean()
  r_mean_v = v_data %>% pull(rating) %>% mean()
  
  calc_df = calc_df %>% 
    mutate(
      r_u_ct = rating_u - r_mean_u,
      r_v_ct = rating_v - r_mean_v,
      r_uv_ct = r_u_ct * r_v_ct,
      r_u_ct_sq = r_u_ct^2,
      r_v_ct_sq = r_v_ct^2
    )
  
  sum_r_uv_ct = calc_df %>% pull(r_uv_ct) %>% sum()
  sum_r_u_ct_sq = calc_df %>% pull(r_u_ct_sq) %>% sum()
  sum_r_v_ct_sq = calc_df %>% pull(r_v_ct_sq) %>% sum()
  
  sim = sum_r_uv_ct / sqrt(sum_r_u_ct_sq * sum_r_v_ct_sq)
  
  return(sim)
}
```

We built a function, `calc_pc_sim`, to calculate the pearson correlation according
to the above formula.

&nbsp;

##### Measure by cosine similarity

Alternatively, $w_{uv}$ can also be defined by cosine similarity:

$$
w_{u v}=\frac{\displaystyle \sum_{m \in M} r_{u m} \cdot r_{v m}}{ \displaystyle \sqrt{\sum_{m \in M}r_{u m}^{2} \cdot \sum_{m \in M}r_{v m}^{2}}}
$$

```{r}
calc_cos_sim = function(u_data, v_data){
  
  calc_df = u_data %>% 
    inner_join(v_data, by = "movie_id", suffix = c("_u", "_v"))
  
  # two users should at least watched 3 same movie
  if (nrow(calc_df) < 3) {
    return(-1)
  }
  
  calc_df = calc_df %>% 
    mutate(
      r_uv = rating_u * rating_v,
      r_u_sq = rating_u^2,
      r_v_sq = rating_v^2
    )
  
  sum_r_uv = calc_df %>% pull(r_uv) %>% sum()
  sum_r_u_sq = calc_df %>% pull(r_u_sq) %>% sum()
  sum_r_v_sq = calc_df %>% pull(r_v_sq) %>% sum()
  
  sim = sum_r_uv / sqrt(sum_r_u_sq * sum_r_v_sq)
  
  return(sim)
}
```


We built the function `calc_cos_sim` to calculate the cosine similarity according
to the above formula.

&nbsp;

### Identify similar users

```{r}
acquire_data = function(user){
  return(ratings %>% filter(user_id == user) %>% pull(data))
}

get_similar = function(user, calc_sim){
  
  user_data = acquire_data(user)
  calc_df = ratings %>% 
    filter(user_id != user) %>%
    rename(v_data = data) %>% 
    mutate(u_data = user_data) %>% 
    mutate(
      sim = map2_dbl(u_data, v_data, calc_sim)
    ) %>% 
    select(-u_data, -v_data) %>% 
    arrange(desc(sim)) %>% 
    head(10)
  
  return(calc_df)
}
```

&nbsp;

### Rating prediction

Once we found users similar to the new user, we can predict the new user's rating
towards a movie $m$ by:

$$
\hat{r}_{u m}=\bar{r}_{u}+\frac{\sum_{v \in S(u, K) \cap N(m)} w_{u v}\left(r_{v m}-\bar{r}_{v}\right)}{\sum_{v \in S(u, K) \cap N(m)}\left|w_{u v}\right|}
$$

- $S(u, K)$: the most similar $K$ users of $u$
- $N(m)$: users who have rated on movie $m$


```{r}
predict_rating = function(user, movie, sim_users){
  
  r_means = ratings_tidy %>% 
    group_by(user_id) %>% 
    summarize(r_mean = mean(rating))
  
  calc_df = 
    ratings_tidy %>% 
    filter(movie_id == movie) %>% 
    inner_join(sim_users, by = "user_id") %>% 
    left_join(r_means, by = "user_id") %>% 
    mutate(
      sim_r_diff = sim * (rating - r_mean),
      abs_sim = abs(sim)
    )
  
  sum_sim_r_diff = calc_df %>% pull(sim_r_diff) %>% sum()
  sum_abs_sim = calc_df %>% pull(abs_sim) %>% sum()
  
  r_u_mean = acquire_data(user)[[1]] %>% pull(rating) %>% mean()
  
  rating = r_u_mean + sum_sim_r_diff/sum_abs_sim
  
  return(rating)
}
```

&nbsp;

# Test & Comparison



### experiment data

```{r}
set.seed(777)
experiment_data = ratings_tidy %>% 
  sample_frac(0.01, weight = user_id)
```

Due to limited computing power, we sample 1% of all ratings data as our experiment data.

&nbsp;

### Measure of performance

To compare the performance of Pearson correlation and cosine similarity, we use RMSE
as the measurement:

$$
\operatorname{RMSE}=\frac{\displaystyle\sqrt{\sum_{(u, m) \in Train}\left(r_{u m}-\hat{r}_{u m}\right)^{2}}}{\mid \text { Test } \mid}
$$

&nbsp;

### Experiment design


```{r}
experiment = function(experiment_data){
  
  test = experiment_data %>% 
    sample_frac(0.2, weight = user_id)

  train = experiment_data %>% 
    anti_join(test)

  ratings_tidy = train

  ratings = ratings_tidy %>% 
    nest(movie_id, rating)
  
  cos_sim_users =
    test %>% 
    select(user_id) %>% 
    distinct() %>% 
    mutate(
      sim_users = map(user_id, ~get_similar(user = .x, calc_sim = calc_cos_sim))
      )
  
  pc_sim_users = 
    test %>% 
    select(user_id) %>% 
    distinct() %>% 
    mutate(
      sim_users = map(user_id, ~get_similar(user = .x, calc_sim = calc_pc_sim))
      )
  
  cos_sum_r_diff_sq_df = 
    test %>%
    left_join(cos_sim_users, by = "user_id") %>% 
    mutate(
      r_pred = pmap_dbl(list(user_id, movie_id, sim_users), predict_rating)
    ) %>% 
    drop_na(r_pred) %>% 
    mutate(
      r_diff_sq = (rating - r_pred)^2
    ) 
  
  cos_sum_r_diff_sq = 
    cos_sum_r_diff_sq_df %>% 
    pull(r_diff_sq) %>% 
    sum(na.rm = TRUE)
  
  pc_sum_r_diff_sq_df = 
    test %>%
    left_join(pc_sim_users, by = "user_id") %>% 
    mutate(
      r_pred = pmap_dbl(list(user_id, movie_id, sim_users), predict_rating)
    ) %>% 
    drop_na(r_pred) %>% 
    mutate(
      r_diff_sq = (rating - r_pred)^2
    ) 
  
  pc_sum_r_diff_sq = 
    pc_sum_r_diff_sq_df %>% 
    pull(r_diff_sq) %>% 
    sum(na.rm = TRUE)
  
  cos_rmse = sqrt(cos_sum_r_diff_sq) / nrow(cos_sum_r_diff_sq_df)
  pc_rmse = sqrt(pc_sum_r_diff_sq) / nrow(pc_sum_r_diff_sq_df)
  
  return(list(cos_rmse, pc_rmse))
}

```

In each round of experiment, for each user, we randomly extract 20% records of 
their ratings as the test dataset, and the remained dataset as train dataset. 
Then, we use the two similarity measures separately to find similar users and 
subsequently predict ratings on corresponding movies in the test dataset. Finally, 
we compute the RMSE between predicted ratings and actual ratings in the test data set,
and get the two RMSE values.

&nbsp;

### Perform experiment

```{r, eval=FALSE}
cos_rmse_list = list()
pc_rmse_list = list()


for (i in 1:30) {
  exp_res = experiment(experiment_data)

  cos_rmse_list = append(cos_rmse_list, exp_res[[1]])
  pc_rmse_list = append(pc_rmse_list, exp_res[[2]])

}

cos_rmse_df = tibble(rmse = cos_rmse_list) %>% unnest()
pc_rmse_df = tibble(rmse = pc_rmse_list) %>% unnest()

rmse_df = bind_rows(cos_rmse_df, pc_rmse_df, .id = "sim_type") %>% 
  mutate(sim_type = if_else(sim_type == 1, "cosine similarity", "pearson correlation"))

rmse_df %>% write_csv("./data/experiment_rmse.csv")

```

We performed 30 times of experiment and recorded RMSE values. This process took
several hours, so we wrote the result into a document. When knitting this file,
we read the result from the document and ignored this chunk.

&nbsp;

### Compare RMSE 

The experiment result reveals that choosing pearson correlation as the measure 
of similarity could yield lower RMSE values. Therefore, in our shiny app, we choose to 
use `calc_pc_sim`.

```{r, eval=FALSE}
rmse_df = read_csv("./data/experiment_rmse.csv") 

rmse_df %>% 
  ggplot(aes(x = sim_type, y = rmse)) %>% 
  geom_violin() %>% 
  labs(x = "Similarity Measurement", y = "RMSE")
```

&nbsp;

# Main function

```{r}
main = function(user){
  
  sim_users = get_similar(user, calc_pc_sim)
  
  movies = sim_users %>% 
    select(user_id) %>% 
    left_join(ratings_tidy, by = "user_id") %>% 
    distinct(movie_id) %>% 
    pull(movie_id) 
  
  calc_df = tibble(
    user_id = user,
    movie_id = movies
  )
  
  watched = acquire_data(user)[[1]] %>% pull(movie_id) %>% as.vector()
  
  rec_movies = 
    calc_df %>% 
    filter(!movie_id %in% watched) %>% 
    mutate(
      r_pred = map2_dbl(user_id, movie_id, ~predict_rating(user = .x, movie = .y, sim_users = sim_users))
    ) %>% 
    arrange(desc(r_pred)) %>% 
    head(10) %>% 
    select(movie_id, pred_rating = r_pred)
  
  return(rec_movies)
}
```

We built the main function to implement the whole process of recommendation. For the
given user, we will find top 10 users similar to him/her, and predicts his/her ratings
on movies similar users have seen yet he/she has not watched. Finally, we select 10
movies with the highest predicted rating to recommend to the user.

&nbsp;

# test

Now, we can use the main function to recommend movies for a random user:

```{r, message=FALSE, warning=FALSE}
set.seed(777)

user = round(runif(1, 1, 610))

movies = read_csv("./data/small/movies.csv") %>% 
  janitor::clean_names() %>% 
  select(-genres)
  
main(user) %>% 
  left_join(movies, by = "movie_id") %>% 
  select(title, pred_rating) %>% 
  kableExtra::kbl(digits = 1) %>% 
  kableExtra::kable_styling(font_size = 12)
```


