---
title: "tag_system_recommendation"
author: "Shaocong Zhang"
date: "11/22/2021"
output: html_document
---

Data wrangling on tags

```{r}
library(tidyverse)
library(tm)
library(ggridges)
library(dplyr)
```

```{r}
# data import
movies = read_csv("../data/small/movies.csv")
ratings = read_csv("../data/small/ratings.csv")
tags = read_csv("../data/small/tags.csv")
```

```{r}
# data wrangling
tags.full = 
  tags %>%
  janitor::clean_names() %>%
  mutate(
    tag = strsplit(tag, " ")
  ) %>%
  unnest() %>% 
  mutate(tag = tolower(tag))

remove.list = read_csv("./stop_words_english.txt", col_names = FALSE) %>% 
  pull("X1") %>% 
  as_vector()

remove.tag = 
tags.full %>% 
  filter(!tag %in% remove.list) %>%
  mutate(
    tag = removePunctuation(tag)
  ) %>% 
  filter(
    !tag == ""
  )
```

Now we get the user tagging behavior data. The dataset contains user id, movie id, tag and time stamp (it represents seconds since midnight Coordinated Universal Time of January 1, 1970, but we will not use this variable in this part). Based on these information, we can construct a personalized recommendation algorithm by:

1. Count the most commonly used tags of the user.

2. For each tag, count the movies that have been labeled the most times.

3. For a user, first find his commonly used tags, and then find the most popular movies with these tags to recommend to the user.

For the above algorithm, the formula of user u's interest in movie i is as follows:

$p(u, i)=\sum_{b} \frac{n_{u, b}}{\log \left(1+n_{b}^{(u)}\right)} \frac{n_{b, i}}{\log \left(1+n_{i}^{(u)}\right)}$

where $B(u)$ is the set of tags that user u has labeled, $B(i)$ is the set of tags that movie i is labeled, $n_{u, b}$ is the number of times that user u has labeled tag b, $n_{b, i}$ is the number of times that movie i has been labeled tag b, $n_b^{(u)}$ records how many different users have used tag b, $n_i^{(u)}$records how many different users have tagged the movie i.

```{r}
# Set up interest function
fc.interest = function(u, i) {
B.u = 
  remove.tag %>% 
  filter(user_id == u)

B.u.distinct = 
  B.u %>% 
  distinct(tag)

remove.tag.distinct = 
  remove.tag %>% 
  select(-movie_id, -timestamp) %>% 
  distinct()

B.i = 
  remove.tag %>% 
  filter(movie_id == i)

B.i.distinct = 
  B.i %>% 
  select(-tag, -timestamp) %>% 
  distinct()

n_ub = c()
n_bi = c()
n_bu = c()
lg.value = c()
n_iu = nrow(B.i.distinct)
value = c()

for (b in 1:nrow(B.u.distinct)) {
  n_ub[b] = sum(B.u$tag == as.character(B.u.distinct[b, 1]))
  n_bi[b] = sum(B.i$tag == as.character(B.u.distinct[b, 1]))
  n_bu[b] = sum(remove.tag.distinct$tag == as.character(B.u.distinct[b, 1]))
  lg.value[b] = log(1 + n_bu[b]) * log(1 + n_iu)
  value[b] = n_ub[b] * n_bi[b] / lg.value[b]
}
return(interest = round(sum(value), 2))
}
```

# Search the movie that was most frequently labeled by the users

Since the dataset we have only contains about 3700 tags labeled by the users. If 

```{r}

```

# Create heat map

```{r}
set.seed(8105)

id.list = unique(remove.tag$user_id)
movie.list = unique(remove.tag$movie_id)

# randomly choose 30 users and 4 movies
id.new = sort(sample(id.list, 5))
movie.new = sort(sample(movie.list, 10))

# construct a new table of the selected id and movie
heat.map = tibble(user_id = sort(rep(id.new, 10)), movie_id = rep(movie.new, 5))

# add insterest value to the table
interest = c()
for (i in 1:nrow(heat.map)) {
  interest[i] = fc.interest(as.numeric(heat.map[i, 1]), as.numeric(heat.map[i, 2]))
}

heat.map = add_column(heat.map, interest)

# build heat map
heat.map %>% 
  ggplot(aes(y = as.factor(user_id), x = as.factor(movie_id), fill = interest)) +
  geom_tile() +
  geom_raster() +
  labs(
    x = "Movie ID",
    y = "User ID"
  )

```









